---
title: "Class 09"
author: "Karis Kim (A16378938)"
date: "2024-04-30"
output:
  pdf_document: default
---

```{r}
candy_url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy <- read.csv(candy_url, row.names = 1)

head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
num_candy <- nrow(candy)
sum(num_candy)
```
> Q2. How many fruity andy types are in the dataset?

```{r}
num_fruity_candy <- sum(candy$fruity)
sum(candy$fruity)
```

> Q3. What is your favorite candy in the dataset and what is it's winpercent value?

```{r}
fav_candy <- "Twix"
fav_winpercent <- candy[fav_candy, ]$winpercent
sum(fav_winpercent)
```

> Q4. What is the winpercent value for **Kit Kat**?

```{r}
kit_kat <- "Kit Kat"
kit_kat_winpercent <- candy[kit_kat, ]$winpercent
sum(kit_kat_winpercent)
```
> Q5. What is the winpercent value for **Tootsie Roll Snack Bars**?

```{r}
tootsie_roll <- "Tootsie Roll Snack Bars"
tootsie_roll_winpercent <- candy[tootsie_roll, ]$winpercent
sum(tootsie_roll_winpercent)
```

Now we use "skimr"

```{r}
library("skimr")

skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

```{r}
skim_output <- skim(candy)
```

```{r}
different_scale <- skim_output$skim$variable[skim_output$skim$type != "logical" & skim_output$skim$type != "integer"]
print(different_scale)
```

> Q7. What do you think a zero and one represent for the candy$chocolate column?

```{r}
chocolate_variable_type <- skim_output$skim$type[skim_output$skim$variable == "chocolate"]
print(chocolate_variable_type)
```

## Histogram of Winpercent Values

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent, main = "Histogram of Winpercent Values", xlab = "Winpercent")
```

> Q9. Is the distribution of winpercent values symmetrical?

Not symmetrical. Right skewed distribution.

> Q10. Is the center of the distribution above or below 50%?

```{r}
winpercent_mean <- mean(candy$winpercent)
sum(winpercent_mean)
```

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate_mean <- mean(candy$winpercent[as.logical(candy$chocolate)])
fruity_mean <- mean(candy$winpercent[as.logical(candy$fruity)])

print(chocolate_mean)
print(fruity_mean)

```

> Q12. Is this difference statistically significant?

```{r}
t_test_result <- t.test(candy$winpercent[as.logical(candy$chocolate)], 
                        candy$winpercent[as.logical(candy$fruity)])
print( t_test_result$p.value)
```

## Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?

```{r}
least_liked_candy <- head(candy[order(candy$winpercent), ], n = 5)
print(least_liked_candy)

library(dplyr)
least_liked_candy_dplyr <- candy %>% arrange(winpercent) %>% head(5)
print(least_liked_candy_dplyr)
```

> Q14. What are the top 5 all-time favorite candy types out of this set?

```{r}
top_favorite_candy <- tail(candy[order(candy$winpercent), ], n = 5)
print(top_favorite_candy)

top_favorite_candy_dplyr <- candy %>% arrange(desc(winpercent)) %>% head(5)
print(top_favorite_candy_dplyr)
```

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_bar(stat = "identity") +
  labs(x = "Winpercent", y = "Candy Type") +
  ggtitle("Candy Ranking Based on Winpercent Values (Sorted)")
```

```{r}
# Set up color vector
my_cols <- rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] <- "chocolate"
my_cols[as.logical(candy$bar)] <- "brown"
my_cols[as.logical(candy$fruity)] <- "red"

# Plot barplot with specified colors
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col(fill = my_cols) +
  labs(x = "Winpercent", y = "Candy Type") +
  ggtitle("Candy Ranking Based on Winpercent Values (Sorted) with Candy Type Colors")

```

> Q17. what is the worst ranked chocolate candy?

Charleston Chew

> Q18. What is the best ranked fruity candy?

Nik L Nip


## Take a Look at a Pricepercent 

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label = rownames(candy)) +
  geom_point(col = my_cols) + 
  geom_text_repel(col = my_cols, size = 3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
candy$win_to_price_ratio <- candy$winpercent / candy$pricepercent
```

```{r}
best_value_candy <- candy[which.max(candy$win_to_price_ratio), "win_to_price_ratio"]
cat("Q19. The candy type offering the most bang for your buck is:", rownames(best_value_candy), "\n")
print(rownames(best_value_candy))
```

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
top_expensive_candies <- head(candy[ord, c("pricepercent", "winpercent")], n = 5)
cat("Q20. The top 5 most expensive candy types in the dataset are:\n")
print(top_expensive_candies)

least_popular_expensive_candy <- top_expensive_candies[which.min(top_expensive_candies$winpercent), ]
cat("The least popular candy among the top 5 most expensive candy types is:", rownames(least_popular_expensive_candy), "\n")
```

# Make a lollipop chart of pricepercent

```{r}
library(ggplot2)

ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col = "gray40") +
  geom_point()
```

## Exploring Correlation Structure

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

```{r}
cij <- cor(candy)
min_corr <- min(cij)
anti_corr_indices <- which(cij == min_corr, arr.ind = TRUE)
anti_corr_variables <- rownames(cij)[anti_corr_indices[1, 1]]
anti_corr_variables2 <- colnames(cij)[anti_corr_indices[1, 2]]
cat("Q22. The two variables that are anti-correlated are:", anti_corr_variables, "and", anti_corr_variables2, "\n")
```

> Q23. Similarly, what two variables are most positively correlated?

```{r}
max_corr <- max(cij)
pos_corr_indices <- which(cij == max_corr, arr.ind = TRUE)

pos_corr_variables <- rownames(cij)[pos_corr_indices[1, 1]]
pos_corr_variables2 <- colnames(cij)[pos_corr_indices[1, 2]]
cat("Q23. The two variables that are most positively correlated are:", pos_corr_variables, "and", pos_corr_variables2, "\n")
```

## Principal Component Analysis 

```{r}
pca <- prcomp(candy, scale = TRUE)

summary(pca)
```

```{r}
plot(pca$x[, 1], pca$x[, 2], 
     xlab = "PC1", ylab = "PC2", 
     main = "PCA Score Plot of PC1 vs PC2")
```


```{r}
plot(pca$x[, 1:2], col = my_cols, pch = 16, 
     xlab = "PC1", ylab = "PC2", 
     main = "PCA Score Plot of PC1 vs PC2")
```

```{r}
my_data <- cbind(candy, pca$x[, 1:3])
library(ggplot2)

p <- ggplot(my_data) + 
  aes(x = PC1, y = PC2, 
      size = winpercent / 100,  
      text = rownames(my_data),
      label = rownames(my_data)) +
  geom_point(col = my_cols)

print(p)
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
ggplotly(p)
```

```{r}
par(mar = c(8, 4, 2, 2))

barplot(pca$rotation[, 1], las = 2, ylab = "PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

```{r}
loadings_pc1 <- pca$rotation[, 1]

strong_positive_loadings_pc1 <- names(loadings_pc1[loadings_pc1 > 0])

cat("Q24. Original variables picked up strongly by PC1 in the positive direction are:", paste(strong_positive_loadings_pc1, collapse = ", "), "\n")
```




