---
title: "Week 6 R Functions"
author: "Karis Kim (A16378938)"
date: "2024-04-21"
output: pdf_document

---

This week we are introducing **R functions** and how to write our own R functions.

Questions to answer:

> Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
plot(1:10)
```


```{r}
# Example input vectors to start with 
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90) 

student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

Follow the guidelines from class

-Write a working snippet of code that solves a simple problem

```{r}
# Straight forward mean()
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)

mean(student1)
```

But... we need to drop the lowest score. First, we need to identify the lowest score.

```{r}
which.min(student1)
```

What I want is to now drop (i.e. exclude) this lowest score from my mean() calculation. 

```{r}
student1[-8]
```

Now we can use the answer from which.min() to return all other elements of the vector. 

```{r}
# This is for our first working snippet
mean( student1[-which.min(student1)] )
```

What about the other example students? Will this work for them?

```{r}
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80) 
mean(student2, na.rm=TRUE)

student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
mean(student3, na.rm=TRUE)
```

Another approach is to mask (i.e. replace) all NA values with zero.

First we need to find the NA elements of the vector? How do we find the NA elements?

```{r}
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80) 
x <- student2

is.na(x)
which( is.na(x) )
```

Now we have identified the NA elements we want to mask them. Replace them with zero?
We will make the NA elements zero.

```{r}
x[is.na(x)] <- 0
x
mean(x)
```

Recall we should drop the lowest score now...

```{r}
x[is.na(x)] <- 0
mean( x[-which.min(x)] )
```

```{r}
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
x <- student3
x[is.na(x)] <- 0
mean( x[-which.min(x)] )
```

## Now We Can Make Our Function

Take the snippet and turn into a function.
Every function has three parts:
1. A Name ( in our case `grade`() )
2. Input Arguments ( a vector of student scores )
3. The Body ( i.e our working snippet )

Using RStudio I will select `Code > Extract Function`

```{r}
grade <- function(x) {
  x[is.na(x)] <- 0
  mean ( x[-which.min(x)] )
}
```

```{r}
grade(student1)
grade(student2)
grade(student3)
```

This looks great! Now we need to add comments to explain this function to our future selves and others who want to use this function. 

```{r}
#' Calculate the average score for a vector of student 
#' dropping the lowest score. 
#' Missing values will be treated as zero.
#' 
#' @param x A numeric vector of homework scores
#'
#' @return An average score
#' @export 
#'
#' @examples
#' student <- (100, NA, 90, 97)
#' grade(Student)
#' 

grade <- function(x) {
  # Mask NA with zero
  # Treat missing values as zero
  x[is.na(x)] <- 0
  # Exclude lowest score from mean
  mean ( x[-which.min(x)] )
}
```

Now we can finally use our function on our real whole class data from the CSV format file: "http://tinyurl.com/gradeinput" 

```{r}
url <- "http://tinyurl.com/gradeinput"
gradebook <- read.csv(url, row.names = 1)
```

```{r}
apply(gradebook, 1, grade)
```

> Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? [3pts]

To answer this question, we will run the apply() function and save the results.

```{r}
results <- apply(gradebook, 1, grade)
sort(results, decreasing = TRUE)
```

```{r}
which.max(results)
```

> Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

```{r}
gradebook
```

```{r}
ave.scores <- apply(gradebook, 2, mean, na.rm=TRUE)
ave.scores
which.min(ave.scores)
```
```{r}
med.scores <- apply(gradebook, 2, median, na.rm=TRUE)
med.scores
which.min(med.scores)
```

```{r}
boxplot(gradebook)
```

> Q4. Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

Are the final results (i.e. average score for each student) correlated with the results (i.e. scores) for the individual homeworks - the gradebook columns.

```{r}
masked.gradebook <- gradebook
masked.gradebook[ is.na(masked.gradebook) ] <- 0
cor(results, masked.gradebook$hw5)
```

```{r}
apply(masked.gradebook, 2, cor, x=results)
```

> Q5. Make sure you save your Quarto document and can click the “Render” (or Rmarkdown”Knit”) button to generate a PDF foramt report without errors. Finally, submit your PDF to gradescope. [1pt]

Knit the document to make a PDF. 